The Correspondence Analysis of Two Mode Networks Revisited

### 2025-02-04 analysis-R1.qmd
plots

# Setup
```{r setup}
      knitr::opts_chunk$set(include = TRUE, 
                       echo = FALSE, 
                       warning = FALSE, 
                       message = FALSE)
      library(conflicted)
      library(cowplot)
      library(ggcorrplot)
      library(ggbump)
      library(ggplot2)
      library(ggpubr)
      library(ggrepel)
      library(here)
      library(factoextra)
      library(igraph)
      library(kableExtra)
      library(networkdata)
      library(pals)
      library(patchwork)
      library(stringr)
      conflicted::conflicts_prefer(dplyr::filter)
```

@@

```{r Faust-Style CA}
   rs <- rowSums(A)
   cs <- colSums(A)
   R <- diag(1/sqrt(rs))
   C <- diag(1/sqrt(cs))
   A.n <- R %*% A %*% C
   ca.res <- svd(A.n)
   N <- sum(A)
   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rs)
   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/cs)
   d1 <- c(d1.r, d1.c)
   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rs)
   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/cs)
   d2 <- c(d2.r, d2.c)
```

```{r Two Mode Eigenvector Scores}
   eig.r <- eigen(A %*% t(A))$vectors
   eig.c <- eigen(t(A) %*% A)$vectors
```

```{r Reflections Bump Charts}
   source(here("Functions", "reflections.R"))
   source(here("Functions", "ref.long.dat.R"))
   evens <- function(x) subset(x, x %% 2 == 0)
   odds <- function(x) subset(x, x %% 2 != 0)
   a <- reflections(A, iter = 26)
   p.e.r <- a$p.r[, evens(1:20)]
   g.e.r <- a$g.r[, evens(1:20)]
   b <- ref.long.dat(a = p.e.r, b = g.e.r, max.iter = 26)
   plot.dat <- b$person
   p <- ggplot(data = filter(plot.dat, n.iter < 18),
               mapping = aes(x = n.iter, y = ref, color = person))
   # geoms
   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
   p <- p + geom_point(size = 5)
   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
                                    aes(x = n.iter, label = person), 
                                    size = 5, nudge_x = -3)
   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 18), 
                              aes(x = n.iter, label = person), 
                              size = 5, nudge_x = 0.01)

@@

```{r Row Schastic Projections Plot}
   D.p <- diag(1/rowSums(A)) #inverse of degree matrix of persons
   P.pg <- D.p %*% A
   rownames(P.pg) <- rownames(A) 
   D.g <- diag(1/colSums(A)) #inverse of degree matrix of groups
   P.gp <- D.g %*% t(A)
   rownames(P.gp) <- colnames(A)
   P.pp <- P.pg %*% P.gp
   P.gg <- P.gp %*% P.pg
   rownames(P.pp) <- substr(colnames(P.pp), start = 1, stop = 3)
   rownames(P.gg) <- substr(colnames(P.gg), start = 1, stop = 3)
   P.p <- ggcorrplot(P.pp, lab = TRUE) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
   P.g <- ggcorrplot(P.gg, lab = TRUE) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
   png(file = here("Plots", "p-norm.png"), width=600, height=500)
   P.p
   dev.off() 
   png(file = here("Plots", "g-norm.png"), width=600, height=500)
   P.g
   dev.off() 
```

```{r Normalized similarity Matrices}
      S.r <- A %*% diag(1/cs) %*% t(A) #degree-normalized similarity matrix for rows
      S.c <- t(A) %*% diag(1/rs) %*% A #degree-normalized similarity matrix for columns
      eig.Sr <- eigen(S.r)
      eig.Sc <- eigen(S.c)
```

```{r CA Eigenvector Plot} 
      eigvec.dat.r <- data.frame(rank = rank(d1.r), 
                             value = as.numeric(d1.r),
                             lab = names(d1.r))
      eigvec.dat.c <- data.frame(rank = rank(d1.c), 
                             value = as.numeric(d1.c),
                             lab = names(d1.c))
      eigval.dat.r <- data.frame(k = 1:nrow(A), value = as.numeric(eig.Sr$values)) %>% 
             mutate(value = value / max(value))
      eigval.dat.c <- data.frame(k = 1:ncol(A), value = as.numeric(eig.Sc$values)) %>% 
             mutate(value = value / max(value))

@@

   # CA Similarity Plot (Groups)
   p <- ggcorrplot(norm.sim.c)
   p <- p + scale_x_discrete(position = "top")
   p <- p + scale_y_discrete(limits = rev)
   b <- p + theme(legend.position = "none",
                 axis.text.x = element_text(hjust = -0.2, size = 10)) 
   # CA Correspondence Plot
   plot.dat <- data.frame(d1, d2, mode = c(rep(1, 18), rep(2, 14)))
   p <- ggplot(plot.dat, aes(y = d2, x = d1, color = factor(mode), label = rownames(plot.dat)))
   p <- p + geom_vline(xintercept = 0, color = "gray")
   p <- p + geom_hline(yintercept = 0, color = "gray")
   p <- p + geom_text_repel(size = 4, max.overlaps = 50)
   p <- p + theme_minimal() 
   p <- p + theme(legend.position = "none",
                  axis.text = element_text(size = 18),
                  axis.title = element_text(size = 20)
                  )
   p <- p + labs(x = "Second Dimension", y = "First Dimension")
   c <- p + scale_color_manual(values = c('red', 'blue'))
   png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=900)
   (a + b) / c
   dev.off()
```

@@

   # Bonacich Similarity Plot (Groups)
   p <- ggcorrplot(norm.sim.c)
   p <- p + scale_x_discrete(position = "top")
   p <- p + scale_y_discrete(limits = rev)
   b <- p + theme(legend.position = "none",
                 axis.text.x = element_text(hjust = -0.2, size = 10)) 
   # Bonacich Correspondence Plot
   eig.d1 <- c(eig.r[, 1], eig.c[, 1])
   eig.d2 <- c(eig.r[, 2], eig.c[, 2])
   plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
   p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
                             label = c(rownames(A), colnames(A))))
   p <- p + geom_vline(xintercept = 0, color = "gray")
   p <- p + geom_hline(yintercept = 0, color = "gray")
   p <- p + geom_text_repel(size = 4, max.overlaps = 50)
   p <- p + theme_minimal() 
   p <- p + theme(legend.position = "none",
                  axis.text = element_text(size = 18),
                  axis.title = element_text(size = 20)
                  )
   p <- p + labs(x = "Second Dimension", y = "First Dimension")
   c <- p + scale_color_manual(values = c('red', 'blue'))
   png(file = here("Plots", "bon-corr-plot.png"), width=1200, height=900)
   (a + b) / c
   dev.off()
```

>>>>

```{r Setup}
      knitr::opts_chunk$set(include = TRUE, 
                       echo = FALSE, 
                       warning = FALSE, 
                       message = FALSE)
      library(conflicted)
      library(cowplot)
      library(ggcorrplot)
      library(ggbump)
      library(ggplot2)
      library(ggpubr)
      library(ggrepel)
      library(here)
      library(factoextra)
      library(igraph)
      library(kableExtra)
      library(networkdata)
      library(pals)
      library(patchwork)
      library(stringr)
      conflicted::conflicts_prefer(dplyr::filter)
      source(here("Functions", "reflections.R"))
      source(here("Functions", "ref.long.dat.R"))
      source(here("Functions", "SimRank.R"))
```

@@

```{r Faust-Style Direct CA}
   ca.res <- svd(diag(1/sqrt(rowSums(A))) %*% A %*% diag(1/sqrt(colSums(A))))
   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rowSums(A))
   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/colSums(A))
   d1 <- c(d1.r, d1.c)
   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rowSums(A))
   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/colSums(A))
   d2 <- c(d2.r, d2.c)
   d3.r <- ca.res$d[4] * ca.res$u[, 4] * sqrt(N/rowSums(A))
   d3.c <- ca.res$d[4] * ca.res$v[, 4] * sqrt(N/colSums(A))
   d3 <- c(d3.r, d3.c)
   d4.r <- ca.res$d[5] * ca.res$u[, 5] * sqrt(N/rowSums(A))
   d4.c <- ca.res$d[5] * ca.res$v[, 5] * sqrt(N/colSums(A))
   d4 <- c(d4.r, d4.c)
```

```{r Dual Projection CA}
   P.pg <- diag(1/rowSums(A)) %*% A
   rownames(P.pg) <- rownames(A) 
   P.gp <- diag(1/colSums(A)) %*% t(A)
   rownames(P.gp) <- colnames(A)
   P.pp <- P.pg %*% P.gp
   P.gg <- P.gp %*% P.pg
   CA.r <- eigen(P.pp)
   CA.c <- eigen(P.gg)
   rownames(CA.r$vectors) <- rownames(A)
   rownames(CA.c$vectors) <- colnames(A)
```

```{r Weighted Similarity CA}
   S.r <- A %*% diag(1/colSums(A)) %*% t(A)
   S.c <- t(A) %*% diag(1/rowSums(A)) %*% A
   SCA.r <- eigen(diag(nrow(A)) - (diag(1/rowSums(A)) %*% S.r))
   SCA.c <- eigen(diag(ncol(A)) - (diag(1/colSums(A)) %*% S.c))
   rownames(SCA.r$vectors) <- rownames(A)
   rownames(SCA.c$vectors) <- colnames(A)
```

```{Two Dual Projection Eigenvector Scores}
   eig.r <- eigen(A %*% t(A))$vectors
   eig.c <- eigen(t(A) %*% A)$vectors
   rownames(eig.r) <- rownames(A)
   rownames(eig.c) <- colnames(A)
```

```{r Reflections Bump Charts}
   evens <- function(x) subset(x, x %% 2 == 0)
   odds <- function(x) subset(x, x %% 2 != 0)
   a <- reflections(A, iter = 26)
   p.e.r <- a$p.r[, evens(1:20)]
   g.e.r <- a$g.r[, evens(1:20)]
   b <- ref.long.dat(a = p.e.r, b = g.e.r, max.iter = 26)
   plot.dat <- b$person
   p <- ggplot(data = filter(plot.dat, n.iter < 18),
               mapping = aes(x = n.iter, y = ref, color = person))
   # geoms
   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
   p <- p + geom_point(size = 5)
   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
                                    aes(x = n.iter, label = person), 
                                    size = 5, nudge_x = -3)
   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 18), 
                              aes(x = n.iter, label = person), 
                              size = 5, nudge_x = 0.01)

@@

```{r Row Stochastic Matrices for CA Projections Plot}
   D.p <- diag(1/rowSums(A)) #inverse of degree matrix of persons
   P.pg <- D.p %*% A
   rownames(P.pg) <- rownames(A) 
   D.g <- diag(1/colSums(A)) #inverse of degree matrix of groups
   P.gp <- D.g %*% t(A)
   rownames(P.gp) <- colnames(A)
   P.pp <- P.pg %*% P.gp
   P.gg <- P.gp %*% P.pg
   rownames(P.pp) <- substr(colnames(P.pp), start = 1, stop = 3)
   rownames(P.gg) <- substr(colnames(P.gg), start = 1, stop = 3)
   P.p <- ggcorrplot(P.pp, lab = TRUE) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
   P.g <- ggcorrplot(P.gg, lab = TRUE) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
   png(file = here("Plots", "p-norm.png"), width=600, height=500)
   P.p
   dev.off() 
   png(file = here("Plots", "g-norm.png"), width=600, height=500)
   P.g
   dev.off() 
```

```{r CA Eigenvector Plot} 
      eigvec.dat.r <- data.frame(rank = rank(d1.r), 
                             value = as.numeric(d1.r),
                             lab = names(d1.r))
      eigvec.dat.c <- data.frame(rank = rank(d1.c), 
                             value = as.numeric(d1.c),
                             lab = names(d1.c))
      eigval.dat.r <- data.frame(k = 1:nrow(A)-1, value = as.numeric(CA.r$values)) %>% 
             mutate(value = value / max(value))
      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, value = as.numeric(CA.c$values)) %>% 
             mutate(value = value / max(value))

@@

      # CA Similarity Plot (Groups)
      p <- ggcorrplot(norm.sim.c)
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      b <- p + theme(legend.position = "none", plot.margin = margin(l = 100), 
                    axis.text.x = element_text(hjust = -0.2, size = 10))

      # CA Correspondence Plot
      plot.dat <- data.frame(a = d3, 
                             b = d1, 
                             mode = c(rep(1, 18), rep(2, 14)))
      p <- ggplot(plot.dat, aes(y = a, x = b, color = factor(mode), label = rownames(plot.dat)))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      p <- p + labs(x = "Second Dimension", y = "First Dimension")
      c <- p + scale_color_manual(values = c('red', 'blue'))
      png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=900)
      (a + b) / c + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off()
```

@@

      # Bonacich Similarity Plot (Groups)
      p <- ggcorrplot(norm.sim.c)
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      b <- p + theme(legend.position = "none", plot.margin = margin(l = 100),
                    axis.text.x = element_text(hjust = -0.2, size = 10))

      # Bonacich Correspondence Plot
      eig.d1 <- c(eig.r[, 1], eig.c[, 1])
      eig.d2 <- c(eig.r[, 2], eig.c[, 2])
      plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
      p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
                                label = c(rownames(A), colnames(A))))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      p <- p + labs(x = "Second Dimension", y = "First Dimension")
      c <- p + scale_color_manual(values = c('red', 'blue'))
      png(file = here("Plots", "bon-corr-plot.png"), width=1200, height=900)
      (a + b) / c + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off()
```

```{r CA versus SimRank}
      res <- SimRank(A)
      Sr <- res$Sr
      Sc <- res$Sc
      s1.r <- eigen(Sr)$vectors[, 2]
      s1.c <- eigen(Sc)$vectors[, 2]
      diag(Sr) <- 0
      diag(Sc) <- 0
      Sr <- Sr[order(d1.r), order(d1.r)]
      Sc <- Sc[order(d1.c), order(d1.c)]
      rownames(Sr) <- substr(colnames(P.pp), start = 1, stop = 3)
      rownames(Sc) <- substr(colnames(P.gg), start = 1, stop = 3)
      a <- ggcorrplot(Sr, lab = TRUE, show.diag = FALSE, lab_size = 3) +
         theme(legend.position = "none", 
               axis.text.x = element_text(hjust = -0.2)) + 
         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
                              low = "white", mid = "yellow", high = "purple")

      b <- ggcorrplot(Sc, lab = TRUE, show.diag = FALSE, lab_size = 3) +
         theme(legend.position = "none", 
               axis.text.x = element_text(hjust = -0.2)) + 
         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
                              low = "white", mid = "yellow", high = "purple")

      dat.r <- data.frame(ca = d1.r,
                          sr = s1.r,
                          lab = rownames(A))
      dat.c <- data.frame(ca = d1.c,
                          sr = s1.c,
                          lab = colnames(A))
      cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)

      c <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
                     label = "lab", repel = TRUE, 
                     add = "reg.line", 
                     add.params = list(color = "gray80", 
                                       linewidth = 0.5, linetype = 2)) +
                     annotate("text", x = -.2, y = .3, 
                              label = paste("r = ", cor), size = 6) 
      cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)

      d <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
                     label = "lab", repel = TRUE, 
                     add = "reg.line",
                     add.params = list(color = "gray80", 
                                       linewidth = 0.5, linetype = 2)) +
                     annotate("text", x = -.2, y = .3, 
                              label = paste("r = ", cor), size = 6) 
      png(file = here("Plots", "simrank-v-ca.png"), width=1000, height=900)
      (a + c + plot_layout(widths = c(2, 1))) / (b + d + plot_layout(widths = c(2, 1))) + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off() 
```
